---
title: "Preprocomb"
author: "Markus Vattulainen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preprocomb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Preprocessing is often the most time-consuming phase in knowledge
discovery and preprocessing transformations interdependent in unexpected
ways. This package helps to make preprocessing faster and more effective. It
provides an S4 framework for creating and testing preprocessing combinations
for classification, clustering and outlier detection. The framework supports
user-defined and domain-specific preprocessors and preprocessing phases.
Default preprocessors can be used for low variance removal, missing value
imputation, scaling, outlier removal, noise smoothing, feature selection and
class imbalance correction.

## Test data

Let's add contaminations to Iris-data to simulate data quality problems:

```{r}
set.seed(1)
testdata <- iris
testdata[,2] <- rnorm(150, testdata[,2], 1) # add noise to the second variable
testdata[sample(1:150,20),3] <- sample(20:40,20, replace=TRUE) # add outliers to the third
testdata$Irrelevant <- runif(150, 0, 1) # add an irrelevant feature
testdata <- testdata[1:110,] # add class imbalance
```

## Interactive mode

In the interactive mode we can apply preprocessing techniques in sequence and
see the end result on classification accuracy. In the below the data is first
scaled and then outliers removed with Orh-algorithm. K-nearest neighbour is 
the default classifier.


```{r, messages=FALSE, warnings=FALSE}
library(preprocomb)
step1 <- prepro(testdata, "basicscale")
step2 <- prepro(step1, "orhoutlier")
step2
```

## Programmatic mode

In the programmatic mode we can search for the best combinations by first setting
the grid of combinations and the evaluating each combination. In the example below
the preprocessing pipeline consists scaling and outlier removal techniques.

```{r, messages=FALSE, warnings=FALSE}
grid <- setgrid(phases=c("scaling", "outliers"), data=testdata)
result <- preprocomb(grid=grid, nholdout=2)
result@allclassification
```

## Default options

The package is intended to be used with domain specific preprocessing phases and 
techniques. There are however a set of default options available. Phases:

- imputation, missing value imputation
- variance, low variance removal
- smoothing, noise smoothing
- scaling, value range scaling 
- outliers, outlier removal
- sampling, class imbalance corrections
- selection, irrelevant feature selection

Each of the phases has two or more preprocessing techniques.
